 Casting multiple shadows: high-dimensional interactive data visualisation with tours and embeddings

### 2020-05-04 paper.Rmd
clarifying tour section

A complementary approach for visualizing structure in high dimensional data is
the tour. A tour is a sequence of projections of a high dimensional dataset
onto a low-dimensional orthonormal basis matrix, that is represented as a
dynamic visualization. The sequence of generated bases are interpolated to form
the tour path, allowing a user to explore the subspace of projections. A grand
tour corresponds to choosing new bases at random, and can give an overview of
the structure in the data. Instead of picking projections at random, a guided
tour can be used to generate a sequence 'interesting' projections as quantified
by an index function. Given the dynamic nature of the tour, user interaction is
important for controlling and exploring the visualisation: the tour has been
used previously by @Wickham2015-cx as tool for exploring statistical model fits
and by @Buja1996-fk for exploring factorial experimental designs.

@@

?? include details of tour here

@@

@@

@@

>>>>

A complementary approach for visualizing structure in high dimensional data is 
the tour. A tour is a sequence of projections of a high dimensional dataset
onto a low-dimensional basis matrix, and is represented as a
animated visualization.Given the dynamic nature of the tour, user interaction is important for controlling and exploring the visualisation: the tour has been
used previously by @Wickham2015-cx as tool for exploring statistical model fits and by @Buja1996-fk for exploring factorial experimental designs.

@@

synthesis of some other techniques like LLE, Isomap and PHATE

@@

### Tours explore the subspace $d$-dimension projections

Here we provide an overview of the mathematical aspects of the tour for
readers who are unfamiliar with the technique before discussing the computational aspects of how `liminal` interfaces with them. The tour
is emphatically not a dimension reduction technique but rather a visualisation technique that is grounded in mathematical theory, that is able to ascertain the shape and global structure of a dataset via inspection
of the subspace generated by the set of low-dimensional projections. In fact, when we use the tour in the case studies, we will first sphere our data via principal components to reduce dimensionality of our input.

Like, when using DR techniques, the tour assumes we have a real data matrix
$X$ consisting of $n$ observations in $p$ dimensions. First, the tour generates 
a  sequence of $p \times d$ orthonormal projection matrices (bases) $A_{t \in \mathbb{N}}$, where $d$ is typically $1$ or $2$. For each pair of orthormal bases $A_{t}$ and $A_{t+1}$ that are generated, the geodesic
path between them is interpolated to form intermediate frames, and giving the sense of continuous movement from one basis to another. The tour is then the continous visualisation of the  projections $Y_{t} = XA_{t}$, that is the projection of $X$ onto $A_{t]$ as the tour path is interpolated between successive bases. A *grand tour* corresponds to choosing new orthoromal bases at random; allowing a user to ascertain structure via exploring the subspace of $d$-dimensional projections. Instead of picking projections at random, a *guided tour* can be used to generate a sequence 'interesting' projections as quantified by an index function. While `liminal` is able to visualise guided tours, our focus is on the grand tour.

@@

The process of generating successive bases and interpolating between them before frames, results in the tour being a dynamic visualisation technique. Generally, if $d$ is one or two dimensions, the results are visualised as
animated density or scatter plots. 
The tour is enhanced by **interactivity**:

- play/pause the current projection
- zooming to find 'interesting' points
- highlighting sets of points

@@

- linking multiple views
